% !TeX root = ../../thesis.tex
\chapter{Predoc research contributions}
\label{ch:myappendix}

The following is a list of papers (both published and in progress) that I have contributed to during my predoc year:

\textbf{Barney I. Potter}, Rebecca Kondor, James Hadfield, John Huddleston, John Barnes, Thomas Rowe, Lizheng Guo, Xiyan Xu, Richard A. Neher, Trevor Bedford, David E. Wentworth, Evolution and rapid spread of a reassortant A(H3N2) virus that predominated the 2017–2018 influenza season. Virus Evolution, Volume 5, Issue 2, July 2019, vez046, https://doi.org/10.1093/ve/vez046

Michael J. Song, \textbf{Barney I. Potter}, Jeff J. Doyle, Jeremy E. Coate. Gene Balance Predicts Transcriptional Responses Immediately Following Ploidy Change in \textit{Arabidopsis thaliana}. The Plant Cell May 2020, 32 (5) 1434-1448; DOI: 10.1105/tpc.19.00832

Kristof Theys, Anna Zhukova, \textbf{Barney I. Potter}, Hanna Guimaraes, Alexander E. Gorbalenya, Ana Tereza Ribeiro de Vasconcelos, Simon Dellicour, Pieter Libin, Ricardo Khouri, Olivier Gascuel, Ana B. Abecasis, Anne-Mieke Vandamme, Guy Baele, Lize Cuypers. Genome-wide diversity of Zika virus : a spatio-temporal perspective on sequence space. \textit{In preparation.}

\textbf{Barney I. Potter}, Marijn Thijssen, N\'{i}dia Sequeira Trov\~{a}o,  Andrea Pineda-Pe\~{n}a, Marijke Reynders, Thomas Mina, Carolina Alvarez, Samad Almini-Bavil-Olyaee, Frederik Nevens, Piet Maes, Philippe Lemey, Marc Van Ranst, Guy Baele, Mahmoud Reza Pourkarim, A comprehensive evolutionary analysis of recent African HBV strains seeding from the heart of Europe. \textit{In preparation.}

%GB: ask Liana how that's going, seems like nothing is happening
%Lassa in Nigeria

%GB: too early to be mentioned here in my opinion
%Will GLUE-BEAST be a paper?

Laura K. Jackson, \textbf{Barney I. Potter}, Sean Schneider, Matthew Fitzgibbon, Kris Blair, Hajirah Farah, Trevor Bedford, Richard Peek Jr., Nina R. Salama. \textit{Helicobacter pylori} diversification during chronic infection within a single host generates sub-populations with distinct phenotypes. \textit{In preparation.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Keep the following \cleardoublepage at the end of this file,
% otherwise \includeonly includes empty pages.
\cleardoublepage

% vim: tw=70 nocindent expandtab foldmethod=marker foldmarker={{{}{,}{}}}
